#### **Title:** CS598-PSL-MovieRecommendation - Final Project
#### **Team members:** gsr2 and balajis2
#### **Date:** 12/7/2020

##### Load the required library files and the dataset

```{r message=FALSE}
library(data.table)
library(dplyr)

MoviesDataURL = paste0("data/","movies.dat")
RatingsDataURL = paste0("data/","ratings.dat")
UserDataURL = paste0("data/","users.dat")

movies = readLines(MoviesDataURL)
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)
movies$Title = iconv(movies$Title, "latin1", "UTF-8")

small_image_url = "https://liangfgithub.github.io/MovieImages/"
movies$image_url = sapply(movies$MovieID, 
                          function(x) paste0(small_image_url, x, '.jpg?raw=true'))
movies$seqno = 1:nrow(movies)

# reading the ratings data
ratings = read.csv(RatingsDataURL, sep = ':', 
                   colClasses = c('integer', 'NULL'), header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')

# Splitting the genres into different Movie genre columns
movie_genre_df = as.data.frame(movies$Genres,stringsAsFactors = FALSE)
movie_genre_df = as.data.frame(tstrsplit(movie_genre_df[,1],"[|]",type.convert = TRUE))
colnames(movie_genre_df) = c("Genre1","Genre2","Genre3","Genre4","Genre5","Genre6")
movie_genre_df[,"MovieID"] = movies$MovieID

```
**For system 1**, we considered two approaches A and B for determining the
popularity of a movie based on the user selected genre

**_For system 1 - Approach A_**

1. Filter all the movies in the dataframe based on the input genre

2. Calculate total number of movies and total number of rating for all 
   the movies in the selected genre

3. Calculate average star rating for each movie (similar to how 
  amazon calculates average rating for each products in their website)
    
  average star rating = (no of 5 star ratings for the movie) * 5 +
                        (no of 4 star ratings for the movie) * 4 +      
                        (no of 3 star ratings for the movie) * 3 +
                        (no of 2 star ratings for the movie) * 2 +
                        (no of 1 star ratings for the movie) * 1 
                        / total no. of ratings for the movie 
                        
4. calculate the weights for each movie as shown below

   weight factor = (total reviews for the movie)/(total number of reviews for     all the movies)
   
5. Multiple the average star rating with the weight factor to generate the score

6. Order the score in increasing order and filter the first 10 records

The above approach ensures that the movies with low number of ratings
but have more 5 stars and 4 stars gets a fair chance in the ranking

```{r}
get_movie_genre_recomm1 <- function(genre_input,movie_genre,ratings) {
  
  if(genre_input == "Children"){
    userSelectedGenre = "Children's"
  }
  else{
    userSelectedGenre = genre_input  
  }
  
  dat.query <- filter(movie_genre, Genre1==userSelectedGenre|Genre2==userSelectedGenre|Genre3==userSelectedGenre|Genre4==userSelectedGenre|Genre5==userSelectedGenre|Genre6==userSelectedGenre)
  MoviesByGenre <- ratings %>% filter(MovieID %in% dat.query$MovieID)
  
  totalmovies = nrow(dat.query)
  totalreviews = nrow(MoviesByGenre)
  
  genre_matrix = matrix(0,totalmovies,11)
  
  colnames(genre_matrix) = c("MovieID","totalratingbygenre","totalratingbyMovie","5star","4star","3star","2star","1star","AvgStarRating","TotalWeight","CalcRank")
  
  genre_matrix[,1] = as.integer(dat.query[,"MovieID"])
  genre_matrix[,2] = as.integer(totalreviews)
  
  for (i in 1:nrow(genre_matrix)){
    tmp = MoviesByGenre %>% filter(MovieID %in% genre_matrix[i,1])
    genre_matrix[i,3] = as.integer(nrow(tmp))
    star.5 = tmp %>% filter(Rating %in% 5)
    star.4 = tmp %>% filter(Rating %in% 4)
    star.3 = tmp %>% filter(Rating %in% 3)
    star.2 = tmp %>% filter(Rating %in% 2)
    star.1 = tmp %>% filter(Rating %in% 1)
    genre_matrix[i,4] = as.integer(nrow(star.5))
    genre_matrix[i,5] = as.integer(nrow(star.4))
    genre_matrix[i,6] = as.integer(nrow(star.3))
    genre_matrix[i,7] = as.integer(nrow(star.2))
    genre_matrix[i,8] = as.integer(nrow(star.1))
    genre_matrix[i,9] = ( (genre_matrix[i,4] * 5) + (genre_matrix[i,5]*4) + (genre_matrix[i,6]*3) + (genre_matrix[i,7]*2) + (genre_matrix[i,8]*1))/genre_matrix[i,3]
    genre_matrix[i,10] = genre_matrix[i,3]/genre_matrix[i,2]
    genre_matrix[i,11] = genre_matrix[i,9] * genre_matrix[i,10]
    
  } 
  
  result = genre_matrix[order(genre_matrix[,11],decreasing = TRUE),]
  
  pred_movie_id = c(result[1:10,1])
  
  return(pred_movie_id)
  
}
genreSelected = "Animation"
recommendedMovieIds = get_movie_genre_recomm1(genreSelected,movie_genre_df,ratings)

print(recommendedMovieIds)
```

**_For system 1 - Approach B (IMDB SCORING)_ **

For approach B we followed the scoring for popular movies similar to how IMDB
calculates the movie score and then ordering for top 10 scores and filter the results.

https://help.imdb.com/article/imdb/track-movies-tv/ratings-faq/G67Y87TFYYP6TWAV#ratings

weighted rating (WR) = (v ÷ (v+m)) × R + (m ÷ (v+m)) × C

Where:

R = average for the movie (mean) = (rating)

v = number of votes for the movie = (votes)

m = minimum votes required to be listed in the Top Rated list 

C = the mean vote across the whole report

This formula provides a true 'Bayesian estimate', which takes into account the number of votes each title has received, minimum votes required to be on the list, and the mean vote for all titles

```{r}
get_movie_genre_recomm2 <- function(genre_input,movie_genre,ratings) {
  
  if(genre_input == "Children"){
    userSelectedGenre = "Children's"
  }
  else{
    userSelectedGenre = genre_input  
  }
  
  dat.query <- filter(movie_genre, Genre1==userSelectedGenre|Genre2==userSelectedGenre|Genre3==userSelectedGenre|Genre4==userSelectedGenre|Genre5==userSelectedGenre|Genre6==userSelectedGenre)
  MoviesByGenre <- ratings %>% filter(MovieID %in% dat.query$MovieID)
  
  totalmovies = nrow(dat.query)
  totalreviews = nrow(MoviesByGenre)
  
  # Calculate minimum number of ratings in the filtered genre
  genre_matrix = matrix(0,totalmovies,6)
  m = quantile(MoviesByGenre$Rating,0.90)
  
  colnames(genre_matrix) = c("MovieID","v","m","R","C","Score")
  genre_matrix[,1] = as.integer(dat.query[,"MovieID"])
  
  for (i in 1:nrow(genre_matrix)){
    tmp = MoviesByGenre %>% filter(MovieID %in% genre_matrix[i,1])
    n = as.integer(nrow(tmp))
    star.5 = tmp %>% filter(Rating %in% 5)
    star.4 = tmp %>% filter(Rating %in% 4)
    star.3 = tmp %>% filter(Rating %in% 3)
    star.2 = tmp %>% filter(Rating %in% 2)
    star.1 = tmp %>% filter(Rating %in% 1)
    # total number of ratings for this movie
    genre_matrix[i,"v"] = n
    # minimum number of ratings needed 
    genre_matrix[i,"m"] = m
    #average rating for the movie
    genre_matrix[i,"R"] = ((nrow(star.5) * 5) + (nrow(star.4) * 4) + (nrow(star.3) * 3) + (nrow(star.2) * 2) + (nrow(star.1) * 1))/n
  }
  
  #mean of average rating for this genre
  genre_matrix[,"C"] = mean(genre_matrix[,"R"])
  
  for (i in 1:nrow(genre_matrix)){
    genre_matrix[i,"Score"] = ((genre_matrix[i,"R"] * genre_matrix[i,"v"]) + (genre_matrix[i,"C"] * genre_matrix[i,"m"]))/(genre_matrix[i,"v"] + genre_matrix[i,"m"]) 
  }  
  
  result = genre_matrix[order(genre_matrix[,"Score"],decreasing = TRUE),]
  
  pred_movie_id = c(result[1:10,1])
  
  return(pred_movie_id)
  
}

genreSelected = "Animation"
recommendedMovieIds = get_movie_genre_recomm2(genreSelected,movie_genre_df,ratings)

print(recommendedMovieIds)

```

